PropOcData2

### 2021-09-09 R/PropOc Data 3D.Rmd
Enrichment Distance Stats

#add region
metaData.df <-read.csv('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/images metadata.csv', header = TRUE)
metaData.df$SampleCode<-as.factor(metaData.df$SampleCode)
metaData.df<-metaData.df %>% select(SampleCode, mouse, Position, size, Region, Zdim, xyDim,xySize)#
propOcMidDensMetaLabel<-metaData.df %>%  merge(propOcMidDensitylabeled, by = "SampleCode")

#include Region
propOcMidDensityStatsRegion<-propOcMidDensMetaLabel %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint,Region, Distance, Relation) %>% summarize(meanValue = mean(Value), sdValue = std.error(Value), medianValue = median(Value))
#exclude region
propOcMidDensityStats<-propOcMidDensMetaLabel %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint,Distance, Relation) %>% summarize(meanValue = mean(Value), sdValue = std.error(Value), medianValue = median(Value))

@@

PlotMeanMidDensityPO<-function(relation){
  ggplot()+
    geom_pointrange(data = propOcMidDensityStats %>% filter(TimePoint == "Day 4", Relation== relation),aes(y=meanValue,ymin= meanValue-(sdValue/2), ymax = meanValue+(sdValue/2), x = Distance,color=Condition),size=1, position = position_dodge(width = .6))+
    #geom_point(data = propOcMidDensityStats %>% filter(Relation== "Pa->Sa"),aes(y=medianValue, x = Distance,color=Condition,shape = "Median"),size=2)+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title ="mean 50 percentile Density")+
    facet_grid(Relation ~ .)+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMeanMidDensityPO("Pa->Sa")
PlotMeanMidDensityPO("Sa->Pa")#this is log exponential
PlotMeanMidDensityPO("Pa->Pa")
PlotMeanMidDensityPO("Sa->Sa")
#this wat combines the data accroding to the focus species
PlotMeanMidDensityPOCenter<-function(){
  ggplot()+
    geom_pointrange(data = propOcMidDensityStats %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Sa->Sa","Sa->Pa")),
    #geom_pointrange(data = propOcMidDensityStats %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Pa->Sa","Pa->Pa")),
                      aes(y=meanValue,ymin= meanValue-(sdValue/2), ymax = meanValue+(sdValue/2), x = Distance,color=Relation),size=1, position = position_dodge(width = .6))+
    #geom_point(data = propOcMidDensityStats %>% filter(Relation== "Pa->Sa"),aes(y=medianValue, x = Distance,color=Condition,shape = "Median"),size=2)+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title ="mean 50 percentile Density")+
    facet_grid(.~Condition )+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMeanMidDensityPOCenter()
PlotMeanMidDensityPOCenterRegion<-function(){
  ggplot()+
    #geom_pointrange(data = propOcMidDensityStatsRegion %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Sa->Sa","Sa->Pa")),
    geom_pointrange(data = propOcMidDensityStatsRegion %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Pa->Sa","Pa->Pa")),
                      aes(y=meanValue,ymin= meanValue-(sdValue/2), ymax = meanValue+(sdValue/2), x = Distance,shape = Region,color=Relation),
    size=1, position = position_dodge(width = .6))+
    #geom_point(data = propOcMidDensityStats %>% filter(Relation== "Pa->Sa"),aes(y=medianValue, x = Distance,color=Condition,shape = "Median"),size=2)+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title ="mean 50 percentile Density")+
    facet_grid(.~Condition )+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMeanMidDensityPOCenterRegion()
#Calculate empty space

@@

#Calculate Enrichment
propOcEnrichment<- propOcMidDensitylabeled %>% group_by(SampleCode,Relation)%>%
  select(Value, Density, Distance, Relation,Sample, Strain,TimePoint, Condition, Culture, Strain,SampleCode) %>% mutate(Enrichment = max(Value)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == Value, 1, 0 )) %>% filter(EnrichmentCh==1) %>% 
  mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max
propOcEnrichment$SampleCode<-as.factor(propOcEnrichment$SampleCode)#make samples a factor

@@

#It's unclear how to deal with the issue of multiple enrichment distances, but I think doing a weighted mean makes the most sense:
pOEnrMetaClean<-pOEnrMeta %>% group_by(SampleCode,Condition,TimePoint, Region, Relation)%>%summarise(WmeanDist =weighted.mean(Distance, Density))
#now calculate the Mean per condition:
pOEnrMetaStatsRegion<-pOEnrMetaClean %>% group_by(Condition,TimePoint, Region, Relation) %>%
  summarise(meanED = mean(WmeanDist),medianED = median(WmeanDist), SDED = sd(WmeanDist), SEED = std.error(WmeanDist),per25 = quantile(WmeanDist, .25), per75 = quantile(WmeanDist, .75))

@@

PlotMeanEnrichment50percentNRegion(pOEnrMetaStats %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"),
                            pOEnrMetaClean %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"))
PlotMedianEnrichment50percentNRegion(pOEnrMetaStats %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"),
                            pOEnrMetaClean %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"))

#original figure 
ggplot()+
    geom_pointrange(data = pOEnrMetaStats,aes(y=meanED,ymax=meanED+SEED,ymin=meanED-SEED, x = Condition),position = position_dodge(width = .5))+
  geom_jitter(data = pOEnrMetaClean,aes(y=WmeanDist,x = Condition),size=1, shape = 1, height = 0, width = .2)+
    labs(y = expression("Enrichment distance ("*mu*"m)")) +
    facet_grid(TimePoint~Relation,scales = "free", space = "free")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 12, margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"))+
    panel_border()

ggplot()+
    geom_pointrange(data = pOEnrMetaStatsRegion,aes(y=meanED,ymax=meanED+SEED,ymin=meanED-SEED, shape = Region,x = Condition),
                    position = position_dodge(width = .5))+
    geom_point(data = pOEnrMetaClean,aes(y=WmeanDist,x = Condition, group = Region),position = position_dodge(width = .5),size=1, shape = 1)+
    labs(y = expression("Enrichment distance ("*mu*"m)")) +
    facet_grid(TimePoint~Relation,scales = "free", space = "free")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 12,margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"))+
    panel_border() 
```

>>>>

#add region
metaData.df <-read.csv('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/images metadata.csv', header = TRUE)
metaData.df$SampleCode<-as.factor(metaData.df$SampleCode)
metaData.df<-metaData.df %>% select(SampleCode, mouse, Position, size, Region, Zdim, xyDim,xySize)#
propOcMidDensMetaLabel<-metaData.df %>%  merge(propOcMidDensitylabeled, by = "SampleCode")
#####
#Calculate Stats
##include Region
propOcMidDensityStatsRegion<-propOcMidDensMetaLabel %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint,Region, Distance, Relation) %>% summarize(meanValue = mean(Value), sdValue = std.error(Value), medianValue = median(Value))
##exclude region
propOcMidDensityStats<-propOcMidDensMetaLabel %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint,Distance, Relation) %>% summarize(meanValue = mean(Value), sdValue = std.error(Value), medianValue = median(Value))

@@

PlotMeanMidDensityPO<-function(relation){
  ggplot()+
    geom_pointrange(data = propOcMidDensityStats %>% filter(TimePoint == "Day 4", Relation== relation),aes(y=meanValue,ymin= meanValue-(sdValue/2), ymax = meanValue+(sdValue/2), x = Distance,color=Condition),size=1, position = position_dodge(width = .6))+
    #geom_point(data = propOcMidDensityStats %>% filter(Relation== "Pa->Sa"),aes(y=medianValue, x = Distance,color=Condition,shape = "Median"),size=2)+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title ="mean 50 percentile Density")+
    facet_grid(Relation ~ .)+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMeanMidDensityPO("Pa->Sa")
PlotMeanMidDensityPO("Sa->Pa")#this is log exponential
PlotMeanMidDensityPO("Pa->Pa")
PlotMeanMidDensityPO("Sa->Sa")
#this wat combines the data accroding to the focus species
PlotMeanMidDensityPOCenter<-function(){
  ggplot()+
    geom_pointrange(data = propOcMidDensityStats %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Sa->Sa","Sa->Pa")),
    #geom_pointrange(data = propOcMidDensityStats %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Pa->Sa","Pa->Pa")),
                      aes(y=meanValue,ymin= meanValue-(sdValue/2), ymax = meanValue+(sdValue/2), x = Distance,color=Relation),size=1, position = position_dodge(width = .6))+
    #geom_point(data = propOcMidDensityStats %>% filter(Relation== "Pa->Sa"),aes(y=medianValue, x = Distance,color=Condition,shape = "Median"),size=2)+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title ="mean 50 percentile Density")+
    facet_grid(.~Condition )+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMeanMidDensityPOCenter()
PlotMeanMidDensityPOCenterRegion<-function(){
  ggplot()+
    #geom_pointrange(data = propOcMidDensityStatsRegion %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Sa->Sa","Sa->Pa")),
    geom_pointrange(data = propOcMidDensityStatsRegion %>% filter(TimePoint == "Day 4", Culture == "co") %>% filter(Relation%in%c("Pa->Sa","Pa->Pa")),
                      aes(y=meanValue,ymin= meanValue-(sdValue/2), ymax = meanValue+(sdValue/2), x = Distance,shape = Region,color=Relation),
    size=1, position = position_dodge(width = .6))+
    #geom_point(data = propOcMidDensityStats %>% filter(Relation== "Pa->Sa"),aes(y=medianValue, x = Distance,color=Condition,shape = "Median"),size=2)+
    #scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    labs(title ="mean 50 percentile Density")+
    facet_grid(.~Condition )+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotMeanMidDensityPOCenterRegion()
#####
#Calculate empty space

@@

#####
#Calculate Enrichment
propOcEnrichment<- propOcMidDensitylabeled %>% group_by(SampleCode,Relation)%>%
  select(Value, Density, Distance, Relation,Sample, Strain,TimePoint, Condition, Culture, Strain,SampleCode) %>% mutate(Enrichment = max(Value)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == Value, 1, 0 )) %>% filter(EnrichmentCh==1) %>% 
  mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max
propOcEnrichment$SampleCode<-as.factor(propOcEnrichment$SampleCode)#make samples a factor

@@

#It's unclear how to deal with the issue of multiple enrichment distances, but I think doing a weighted mean makes the most sense:
#THIS is the one used for stats
pOEnrMetaClean<-pOEnrMeta %>% group_by(SampleCode,Condition,TimePoint, Region, Relation)%>%summarise(WmeanDist =weighted.mean(Distance, Density))
pOEnrMetaClean$CondRegion <- paste0(pOEnrMetaClean$Condition, "-",pOEnrMetaClean$Region)
#####
#Stats
PaSaEnDi.Aov <- aov(WmeanDist ~ Condition, data = pOEnrMetaClean%>%
                     filter(Relation == "Pa->Sa",TimePoint == "Day 4"))
SaPaEnDi.Aov <- aov(WmeanDist ~ Condition, data = pOEnrMetaClean%>%
                       filter(Relation == "Sa->Pa",TimePoint == "Day 4"))

PaSaEnDiRgn.Aov <- aov(WmeanDist ~ CondRegion, data = pOEnrMetaClean%>%
                       filter(Relation == "Pa->Sa",TimePoint == "Day 4"))
SapaEnDiRgn.Aov <- aov(WmeanDist ~ CondRegion, data = pOEnrMetaClean%>%
                       filter(Relation == "Sa->Pa",TimePoint == "Day 4"))
# Summary of the analysis #only showing significant values
summary(PaSaEnDi.Aov)#*p = 0.03
TukeyHSD(PaSaEnDi.Aov)#p wt~pqsL = 0.03
summary(SaPaEnDi.Aov)#nope
summary(PaSaEnDiRgn.Aov)#*p = 0.08
TukeyHSD(PaSaEnDiRgn.Aov)#p wt-edge~pqsL-core = 0.04
summary(SapaEnDiRgn.Aov)#nope!

#now calculate the Mean per condition:
pOEnrMetaStatsRegion<-pOEnrMetaClean %>% group_by(Condition,TimePoint, Region, Relation) %>%
  summarise(meanED = mean(WmeanDist),medianED = median(WmeanDist), SDED = sd(WmeanDist), SEED = std.error(WmeanDist),per25 = quantile(WmeanDist, .25), per75 = quantile(WmeanDist, .75))

@@

PlotMeanEnrichment50percentNRegion(pOEnrMetaStats %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"),
                            pOEnrMetaClean %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"))
PlotMedianEnrichment50percentNRegion(pOEnrMetaStats %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"),
                            pOEnrMetaClean %>% filter(TimePoint == "Day 4", Relation == "Pa->Sa", TimePoint == "Day 4"))
#####
#original figures
##no region
ggplot()+
    geom_pointrange(data = pOEnrMetaStats,aes(y=meanED,ymax=meanED+SEED,ymin=meanED-SEED, x = Condition),position = position_dodge(width = .5))+
  geom_jitter(data = pOEnrMetaClean,aes(y=WmeanDist,x = Condition),size=1, shape = 1, height = 0, width = .2)+
    labs(y = expression("Enrichment distance ("*mu*"m)")) +
    facet_grid(TimePoint~Relation,scales = "free", space = "free")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 12, margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"))+
    panel_border() 
##include Region
ggplot()+
    geom_pointrange(data = pOEnrMetaStatsRegion,aes(y=meanED,ymax=meanED+SEED,ymin=meanED-SEED, shape = Region,x = Condition),
                    position = position_dodge(width = .5))+
    geom_point(data = pOEnrMetaClean,aes(y=WmeanDist,x = Condition, group = Region),position = position_dodge(width = .5),size=1, shape = 1)+
    labs(y = expression("Enrichment distance ("*mu*"m)")) +
    facet_grid(TimePoint~Relation,scales = "free", space = "free")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5),
          strip.text = element_text(size = 12,margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=.5, linetype="solid"))+
    panel_border() 
```
